---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dibble

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/UchidaMizuki/dibble/branch/main/graph/badge.svg)](https://app.codecov.io/gh/UchidaMizuki/dibble?branch=main)
[![R-CMD-check](https://github.com/UchidaMizuki/dibble/workflows/R-CMD-check/badge.svg)](https://github.com/UchidaMizuki/dibble/actions)
<!-- badges: end -->

A 'dibble' (derived from 'dimensional tibble') is a data frame consisting of arrays with named dimensions (known as data cubes).
The columns of the dibbles are classified into dimensions or measures, and the operations on the measures are broadcasted.

## Installation

``` r
# the released version from CRAN:
install.packages("dibble")

# the development version from GitHub:
# install.packages("devtools")
devtools::install_github("UchidaMizuki/dibble")
```

## Examples

```{r,message=FALSE}
library(dibble)
library(dplyr)
```

### Broadcasting
```{r}
arr1 <- array(1:6, c(2, 3),
              list(axis1 = letters[1:2],
                   axis2 = letters[1:3]))
arr2 <- array(1:20, c(4, 5),
              list(axis2 = letters[1:4],
                   axis3 = letters[1:5]))

try(arr1 * arr2)

ddf1 <- as_dibble(arr1)
ddf2 <- as_dibble(arr2)

ddf1 * ddf2

# You can use broadcast() to suppress the warnings.
broadcast(ddf1 * ddf2,
          dim_names = c("axis1", "axis2", "axis3"))
```

### How to build a dibble

#### From a data.frame

```{r}
df <- expand.grid(axis1 = letters[1:2],
                  axis2 = letters[1:3]) %>% 
  as_tibble() %>% 
  mutate(value1 = row_number(),
         value2 = value1 * 2)

ddf <- df %>% 
  dibble_by(axis1, axis2)

df
ddf

# You can access the measures from the dibble with `$`.
ddf$value1
```

#### From an array with named dimensions

```{r}
arr <- array(1:6, 2:3,
             list(axis1 = letters[1:2],
                  axis2 = letters[1:3]))

ddf <- as_dibble(arr)

arr
ddf
```

#### From a vector

```{r}
ddf <- dibble(value = 1:6,
              .dim_names = list(axis1 = letters[1:2],
                                axis2 = letters[1:3]))
ddf
ddf$value
```

```{r}
ddf <- broadcast(1:6,
                 list(axis1 = letters[1:2],
                      axis2 = letters[1:3]))

ddf
```
