---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dibble

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/UchidaMizuki/dibble/branch/main/graph/badge.svg)](https://app.codecov.io/gh/UchidaMizuki/dibble?branch=main)
[![R-CMD-check](https://github.com/UchidaMizuki/dibble/workflows/R-CMD-check/badge.svg)](https://github.com/UchidaMizuki/dibble/actions)
<!-- badges: end -->

A 'dibble' is a data frame consisting of arrays with named dimensions (known as data cubes).
The columns of the dibbles are classified into dimensions or measures, and the operations on the measures are broadcasted.

## Installation

``` r
# the released version from CRAN:
install.packages("dibble")

# the development version from GitHub:
# install.packages("devtools")
devtools::install_github("UchidaMizuki/dibble")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r,message=FALSE}
library(dibble)
library(dplyr)
```

A dibble can be created from a data frame using `dibble_by()` function.

```{r}
ddf <- tidyr::expand_grid(axis1 = 1:2,
                          axis2 = 1:3) %>% 
  mutate(value = row_number()) %>% 
  dibble_by(axis1, axis2)

ddf
```

You can access the measures from the dibble with `$`.

```{r}
ddf$value
```

If the axes of one dibble encompasses the axes of another, `mutate()` allows for dimension-preserving broadcasting.

```{r}
ddf1 <- tidyr::expand_grid(axis1 = 1:2,
                           axis2 = 1:3) %>% 
  mutate(value = row_number()) %>% 
  dibble_by(axis1, axis2)

ddf2 <- tidyr::expand_grid(axis2 = 1:4) %>% 
  mutate(value = row_number()) %>% 
  dibble_by(axis2)

ddf1 %>% 
  mutate(value = value + ddf2$value)
```

`dibble()` allows broadcasting based on the union set of axes of multiple dibbles (or measures).

```{r}
ddf1 <- tidyr::expand_grid(axis1 = 1:2,
                           axis2 = 1:3) %>% 
  mutate(value = row_number()) %>% 
  dibble_by(axis1, axis2)

ddf2 <- tidyr::expand_grid(axis2 = 1:4,
                           axis3 = 1:2) %>% 
  mutate(value = row_number()) %>% 
  dibble_by(axis2, axis3)

dibble(value = ddf1$value * ddf2$value)
```
